The language of generalization: Probability, vagueness, and interaction

### 2018-02-13 paper/comgen-paper.Rmd
change out bda figure; tweak formatting

author: 
  - name          : "Michael Henry Tessler"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305"
    email         : "mtessler@stanford.edu"
  - name          : "Noah D. Goodman"
    affiliation   : "1"

@@

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

@@

Despite its ubiquity and relative morphosyntactic simplicity, there have been few formal models of the language of generalization and none that have made precise, quantitative predictions that match human intuitions. 
The major challenge in formalizing generalization in language is determining what makes the statements true or false.
When we hear "Dogs bark", it sounds like a universal statement, as in "All dogs bark".
But the property does not apply to all dogs: Basenji dogs are barkless,
<!-- ^[ -->
<!-- http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html -->
<!-- ],  -->
for example.
The generalization could signal something weaker like "Most dogs bark".
However, *most* mosquitos do not carry malaria (in fact, only a tiny fraction do), but "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only *quantifier* that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria").
But weakening the generalization to mean *some* is too permissive: Some robins are female (in fact, half of them are), but endorsing "Robins are female" seems odd.
Even more perplexing: Those very same female robins lay eggs, and "Robins lay eggs" is a perfectly fine thing to say.

@@

The language of generalization is easy to spot when a property is predicated of a category: "Ks F" (e.g., "Dogs bark").
This is often called *generic language* or *generics* [@Carlson1977].
But generalization can also manifest when describing an individual (*habitual language*; e.g., "John smokes").
Verbs like *causes* or *makes* also seem to convey generalization, in this case, about an instance of a causal event (*causal language*; e.g., "Staring at the sun makes you go blind"). ^[
  We adopt this somewhat idiosyncratic categorization in order to emphasize the different semantic types predicated. 
  Some authors refer to all of these types as *generics*.
  Still, however, the primary "generics" investigated in the literature have been generalizations about categories ("Ks F").
]
Psychologists, linguists, and philosophers have long studied the language of generalization, as it appears very simple (e.g., syntactically) yet its meaning is difficult to formalize.^[
  We take up the problem of formalizing a *meaning* for the language of generalization as opposed to the problem of *identifying* generalizations. 
  Syntax alone is neither necessary nor sufficient for a listener to know that the sentence conveys a generalization  (e.g., indefinite singulars can encode generalizations: "A dog has four legs" and bare plurals may not: "Dogs are on my front lawn").
  Our analysis thus begins once a sentence has been disambiguated as conveying a generalization. 
]

>>>>

author: 
  - name          : "Michael Henry Tessler"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305"
    email         : "mhtessler@stanford.edu"
  - name          : "Noah D. Goodman"
    affiliation   : "1"

@@

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
toc               : yes

@@

Despite its ubiquity and relative morphosyntactic simplicity, the language of generalization has been addressed by few formal models of and none that have made precise, quantitative predictions that match human intuitions. 
The major challenge in formalizing generalization in language is determining what makes the statements true or false.
When we hear "Dogs bark", it sounds like a universal statement, as in "All dogs bark".
But the property does not apply to all dogs: Basenji dogs are barkless,
<!-- ^[ -->
<!-- http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html -->
<!-- ],  -->
for example.
The generalization could signal something weaker like "Most dogs bark".
However, *most* mosquitos do not carry malaria (in fact, only a tiny fraction do), but "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only *quantifier* that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria").
But weakening the generalization to mean *some* is too permissive: Some robins are female (in fact, half of them are), but endorsing "Robins are female" seems odd.
Even more perplexing: Those very same female robins lay eggs, and "Robins lay eggs" is a perfectly fine thing to say.

@@

The language of generalization is easy to spot when a property, that could project onto an individual, is predicated of a category (e.g., "Dogs are friendly"; *barks* could apply to an individual as in "This dog is friendly").
This is often called *generic language* or *generics* [@Carlson1977].
Generic statements can be contrasted with kind-denoting statements which use properties that can only apply to the category as a whole (e.g., "Dinosaurs are extinct"; *extinct* is not a property that could apply to an individual).
Generalization can also manifest when describing an individual (*habitual language*; e.g., "John smokes").
Verbs like *causes* or *makes* also seem to convey generalization, in this case, about an instance of a causal event (*causal language*; e.g., "Staring at the sun makes you go blind"). ^[
  We adopt this somewhat idiosyncratic categorization in order to emphasize the different semantic types predicated. 
  Some authors refer to all of these types as *generics*.
  Still, however, the primary "generics" investigated in the literature have been generalizations about categories ("Ks F").
]
Psychologists, linguists, and philosophers have long studied the language of generalization, as it appears very simple (e.g., syntactically) yet its meaning is difficult to formalize.^[
  We take up the problem of formalizing a *meaning* for the language of generalization as opposed to the problem of *identifying* generalizations. 
  Syntax alone is neither necessary nor sufficient for a listener to know that the sentence conveys a generalization  (e.g., indefinite singulars can encode generalizations: "A dog has four legs" and bare plurals may not: "Dogs are on my front lawn").
  Our analysis thus begins once a sentence has been disambiguated as conveying a generalization. 
]
